Generalization Errors of Hard Parameter Sharing: High-dimensional Asymptotics and Implications

### 2020-09-27 arxiv/diff_cov.tex
sec B wip
wip

When the covariates across tasks are different, we can no longer obtain an optimal solution for minimizing equation \eqref{eq_mtl} as in Claim \ref{lem_exp_opt}.
This is because the minimization problem is non-convex with respect to $A$ and $B$.
Therefore, we consider an arbitrary local minimum of equation \eqref{eq_mtl} and provide generalization bounds that scale with key properties of task data such as sample sizes and covariate shifts.
We focus on the two-task case to better understand the impact of having different sample sizes and different covariates in hard parameter sharing.
We will develop the key technical results in this section, and describe their implications in the next section.

@@

\begin{theorem}\label{thm_main_RMT}
	Suppose that $X_1$ and $X_2$ are both random matrices that satisfy Assumption \ref{assume_rm}.
	With high probability over the randomness of $X_1$ and $X_2$, we have the following limits:
	\begin{enumerate}
		\item[i)]  The variance equation converges to the following asymptotic limit
			\begin{align}\label{lem_cov_shift_eq}
				\bigabs{\bigtr{\Sigma^{(2)} \bigbrace{(n_1 + n_2){\hat{\Sigma}^{-1}} - (a_1 \Sigma^{(1)} + a_2 \Sigma^{(2)})^{-1}}}} \lesssim p^{-c_{\varphi}}\norm{\Sigma^{(2)}},
			\end{align}
			where $a_1$ and $a_2$ are the solutions of the following two equations:
			\begin{align}
				a_1 + a_2 = 1- \frac{p}{n_1 + n_2},\quad a_1 + \frac1{n_1 + n_2}\cdot \bigbrace{\sum_{i=1}^p \frac{\lambda_i^2 a_1}{\lambda_i^2 a_1 + a_2}} = \frac{n_1}{n_1 + n_2}. \label{eq_a12extra}
			\end{align}
		\item[ii)] For any fixed vector $w\in\real^p$ with unit norm, the bias equation converges to the following asymptotic limit
			\begin{align}\label{lem_cov_derv_eq}
				\bigabs{w^{\top} \Sigma^{(1)} \bigbrace{(n_1+n_2)^2\hat{\Sigma}^{-1} \Sigma^{(2)} \hat{\Sigma}^{-1} - {\Sigma^{(2)}}^{-1/2} V {\frac{a_3 \Lambda^2 + (a_4 + 1)\id}{(a_1 \Lambda^2 + a_2\id)^2}} V^{\top} {\Sigma^{(2)}}^{-1/2}} \Sigma^{(1)} w} \lesssim p^{-c_{\varphi}},
			\end{align}
				where $a_{3}$ and $a_4$ are the solutions of the following two equations % with $b_k = \frac1{p}\sum_{i=1}^p \frac{\lambda_i^{2k}} {(\lambda_i^2 a_1 + a_2)^2}$, for $k = 0, 1, 2$:
			\begin{align}\label{eq_a34extra}
				a_3 + a_4 = \frac{1}{n_1 + n_2}\sum_{i=1}^p \frac{1}{\lambda_i^2 a_1 + a_2},
				a_3 + \frac{1}{n_1 + n_2} \sum_{i=1}^p \frac{\lambda_i^2 (a_1 a_4 - a_2 a_3)}{(\lambda_i^2 a_1 + a_2)^2} = \frac{1}{n_1 + n_2} \sum_{i=1}^p \frac{\lambda_i^2 a_1}{(\lambda_i^2 a_1 + a_2)^{2}}.
%				\left(\frac{\rho_1}{a_1^{2}} -  b_2  \right)\cdot  a_3 -  b_1 \cdot  a_4 = b_1,\quad \left(\frac{\rho_2}{a_2^{2}}-  b_0\right)\cdot  a_4 - b_1 \cdot  a_3
%				= b_0.
			\end{align}
	\end{enumerate}
\end{theorem}
\noindent{\textbf{Remark.}} Our result extends a well-known result on the inverse of sample covariance matrices.
In particular, by setting $X_1 = 0$, we obtain the second result in Fact \ref{lem_minv}.
When $X_1=0$, it is not hard to solve $a_1 = 0$ and $a_2 = (n_2-p) / n_2$ from equation \eqref{eq_a12extra}, and our result reduces to the case of a single sample covariance matrix.
For the single-task case of a multivariate Gaussian random matrix, the result follows from the classical result for the mean of inverse Wishart distribution \cite{anderson1958introduction}.
For a non-Gaussian random matrix, the result can be obtained using the Stieltjes transform method (cf. Lemma 3.11 of \citet{bai2009spectral}).

By combining the bias and variance asymptotics, we obtain the generalization error of $\hat{\beta}_2^{\MTL}$.
Since the proof is similar to the same covariates setting, we describe the key differences and omit the rest of the details.
First, we will rescale $X_1$ by $A_1 / A_2$ before appling Theorem \ref{thm_main_RMT}.
Second, we approximate ${X^{(1)}}^{\top}X^{(1)}$ with $n_1 \Sigma^{(1)}$ in equation \eqref{eq_diff_cov} using Fact \ref{lem_minv}.
For the rest of this section, we present the main ideas in the proof of Theorem \ref{thm_main_RMT}.
%We will give a simple (although not totally rigorous) derivation of the equations \eqref{eq_a12extra} and \eqref{eq_a34extra}.
The full proof can be found in Section \ref{sec_maintools}.

>>>>

When the covariates across tasks are different, we can no longer obtain an optimal solution for minimizing equation \eqref{eq_mtl} as in Claim \ref{lem_exp_opt}.
This is because the minimization problem is non-convex with respect to $A$ and $B$.
Therefore, we consider an arbitrary local minimum of equation \eqref{eq_mtl} and provide generalization bounds that scale with key properties of task data such as sample sizes and covariate shifts.
We focus on the two-task case to better understand the impact of having different sample sizes and different covariates in hard parameter sharing.

@@

\begin{theorem}\label{thm_main_RMT}
	Suppose that $X_1$ and $X_2$ are both random matrices that satisfy Assumption \ref{assume_rm}.
	Assume further that: a) the singular values of $M$ are all greater than $\tau$ and less than $1/\tau$; b) the sample size of task two is greater than $(1 + \tau) p$.
	With high probability over the randomness of $X_1$ and $X_2$, we have the following limits:
	\begin{enumerate}
		\item[i)]  The variance equation converges to the following asymptotic limit
			\begin{align}\label{lem_cov_shift_eq}
				\bigabs{\bigtr{\Sigma^{(2)} \bigbrace{(n_1 + n_2){\hat{\Sigma}^{-1}} - (a_1 \Sigma^{(1)} + a_2 \Sigma^{(2)})^{-1}}}} \lesssim p^{-c_{\varphi}}\norm{\Sigma^{(2)}},
			\end{align}
			where $a_1$ and $a_2$ are the solutions of the following two equations:
			\begin{align}
				a_1 + a_2 = 1- \frac{p}{n_1 + n_2},\quad a_1 + \frac1{n_1 + n_2}\cdot \bigbrace{\sum_{i=1}^p \frac{\lambda_i^2 a_1}{\lambda_i^2 a_1 + a_2}} = \frac{n_1}{n_1 + n_2}. \label{eq_a12extra}
			\end{align}
		\item[ii)] For any fixed vector $w\in\real^p$ with unit norm, the bias equation converges to the following asymptotic limit
			\begin{align}\label{lem_cov_derv_eq}
				\bigabs{w^{\top} \Sigma^{(1)} \bigbrace{(n_1+n_2)^2\hat{\Sigma}^{-1} \Sigma^{(2)} \hat{\Sigma}^{-1} - {\Sigma^{(2)}}^{-1/2} V {\frac{a_3 \Lambda^2 + (a_4 + 1)\id}{(a_1 \Lambda^2 + a_2\id)^2}} V^{\top} {\Sigma^{(2)}}^{-1/2}} \Sigma^{(1)} w} \lesssim p^{-c_{\varphi}},
			\end{align}
				where $a_{3}$ and $a_4$ are the solutions of the following two equations % with $b_k = \frac1{p}\sum_{i=1}^p \frac{\lambda_i^{2k}} {(\lambda_i^2 a_1 + a_2)^2}$, for $k = 0, 1, 2$:
			\begin{align}\label{eq_a34extra}
				a_3 + a_4 = \frac{1}{n_1 + n_2}\sum_{i=1}^p \frac{1}{\lambda_i^2 a_1 + a_2},
				a_3 + \frac{1}{n_1 + n_2} \sum_{i=1}^p \frac{\lambda_i^2 (a_1 a_4 - a_2 a_3)}{(\lambda_i^2 a_1 + a_2)^2} = \frac{1}{n_1 + n_2} \sum_{i=1}^p \frac{\lambda_i^2 a_1}{(\lambda_i^2 a_1 + a_2)^{2}}.
%				\left(\frac{\rho_1}{a_1^{2}} -  b_2  \right)\cdot  a_3 -  b_1 \cdot  a_4 = b_1,\quad \left(\frac{\rho_2}{a_2^{2}}-  b_0\right)\cdot  a_4 - b_1 \cdot  a_3
%				= b_0.
			\end{align}
	\end{enumerate}
\end{theorem}
\noindent{\textbf{Remark.}} Our result extends a well-known result on the inverse of sample covariance matrices.
In particular, by setting $X_1 = 0$, we obtain the second result in Fact \ref{lem_minv}.
When $X_1=0$, it is not hard to solve $a_1 = 0$ and $a_2 = (n_2-p) / n_2$ from equation \eqref{eq_a12extra}, and our result reduces to the case of a single sample covariance matrix.
For a multivariate Gaussian random matrix, the result follows from the classical result for the mean of inverse Wishart distribution \cite{anderson1958introduction}.
For a non-Gaussian random matrix, the result can be obtained using the Stieltjes transform method (cf. Lemma 3.11 of \citet{bai2009spectral}).

By combining the bias and variance asymptotics, we obtain the generalization error of $\hat{\beta}_2^{\MTL}$.
Since the proof is similar to the same covariates setting, we describe the key differences and omit the rest of the details.
First, we will rescale $X_1$ by $A_1 / A_2$ before applying Theorem \ref{thm_main_RMT}.
Second, we approximate ${X^{(1)}}^{\top}X^{(1)}$ with $n_1 \Sigma^{(1)}$ in equation \eqref{eq_diff_cov} using Fact \ref{lem_minv}.
For the rest of this section, we present the main ideas in the proof of Theorem \ref{thm_main_RMT}.
%We will give a simple (although not totally rigorous) derivation of the equations \eqref{eq_a12extra} and \eqref{eq_a34extra}.
The full proof can be found in Section \ref{sec_maintools}.

### 2020-09-27 arxiv/examples.tex
wip

%We provide precise explanations to the phenomena of negative transfer in multi-task learning.
%We provide tight bounds on the bias and variance of the multi-task estimator for two tasks.
%The results in Section \ref{sec_general} show when multi-task learning outperforms single-task learning for general task covariance matrices and ground truth model parameters.
%In particular, the general conditions depend on specific properties of task data such as the covariate shift matrix.
%However, the conditions given by the limiting bias and variance of multi-task learning can be difficult to interpret.
%The goal of this section is to address this issue.
%We interpret the limiting bias and variance equations in an isotropic covariance setting that is described below and a simplified covariate-shifted setting.
%We show that three properties of task data affect the performance of multi-task learning as follows.
%(i) \textit{Task similarity}: We explain the phenomenon of negative transfer precisely as tasks become more different.
%(ii) \textit{Sample ratio}: We further explain a curious phenomenon where increasing the sample ratio between the source and target task helps initially, but hurts eventually.
%(iii) \textit{Covariate shift}: Finally, we show that as the sample size of the source task increases, we show that the covariate shift worsens the performance of the multi-task estimator.
%Finally, we provide a extend our results from two tasks to many tasks with the same features.
%We explain from three perspectives, including \textit{task similarity}, \textit{sample size} and \textit{covariate shift}.
%We show how negative transfer occurs by varying task similarity or sample size.
%Then we show that when source task sample size becomes large, covariate shift causes more negative effects.
We illustrate our main results in simplified settings.
We explain several empirical phenomena that have often been observed when hard parameter sharing approaches are used.

%	The labels are $Y_i = X_i\beta_i + \varepsilon_i$, where $\e_i$ consists of i.i.d. entries with mean zero and variance $\sigma^2$.
%	For our purpose, it is enough to think of the order of $d$ being $1/\sqrt{p}$ and $pd^2/\sigma^2$ being constant.
%The more precise conditions on the relations between $d^2$, $\sigma^2$ and $\kappa^2$ are given in  \eqref{choiceofpara}.
%	We assume that all the random variables have subexponential decay, while keeping in mind that our results can be applied under weaker moments assumptions as shown in Appendix \ref{sec_maintools}.

@@

@@

	Imagine that we have two tasks and the covariates of both tasks have isotropic covariance matrices.
%	Each task has sample size $n_1 = \rho_1 \cdot p$ and $n_2 = \rho_2 \cdot p$.
	%And $X_1\in\real^{n_1\times p}, X_2\in\real^{n_2\times p}$ denote the covariates of the two tasks, respectively.
	For the ground truth linear model, we consider a random-effects model where the two tasks differ by an independent Gaussian vector.
	Specially, for task two, we assume that its model parameter, denoted by $\beta_2$, consists of i.i.d. Gaussian entries with mean zero and variance $\kappa^2$.
	For task one, we assume that its model parameter, denoted by $\beta_1$, is equal to $\beta_2$ plus i.i.d. Gaussian entries with mean $0$ and variance $d^2$.
	Therefore, the two tasks have different ground truth models. In particular, the distance between $\beta_1$ and $\beta_2$ roughly scales as $d^2 \cdot p$.

%\begin{example}[Varying model distance]
%\end{example}

Using Theorem \ref{thm_main_RMT}, %our techniques described later in Section \ref{sec_general}, 
we can derive the limiting prediction loss of the hard parameter sharing estimator.
Without loss of generality, we describe the result for task two and remark that a similar result for task one holds.

\begin{proposition}[Derived from Theorem \ref{thm_main_RMT}]\label{cor_MTL_loss}
	Assume that the isotropic setting described above holds, and that there exists a small constant $c_0>0$ such that the signal-to-noise ratios satisfy that
\be\label{choiceofpara0}
p^{ c_0} \le \frac{p\kappa^2}{\sigma^2 }  \le p^{1-c_0},\quad  \frac{pd^2}{\sigma^2}=\OO(1).
\ee
	Let $B, W_1, W_2$ be the global minimum of equation \eqref{eq_mtl}. Then with high probability, the prediction loss of the hard parameter sharing estimator for task two satisfies:
	\begin{align}\label{cor_MTL_error}
	%-\left[1- \left( 1-\frac{1}{\sqrt{\rho_1}}\right)^4\right] pd^2\cdot \frac{\rho_1^2 (\rho_1+\rho_2)}{(\rho_1 + \rho_2 - 1)^3} +\OO(p^{-c}\sigma^2)  \le 
	 & \left|L(\hat{\beta}_2^{\MTL}) -\frac{\sigma^2}{\rho_1+\rho_2-1} - pd^2\cdot \frac{\rho_1^2 (\rho_1+\rho_2)}{(\rho_1 + \rho_2 - 1)^3}\right|  \le  \left[\left( 1+\frac{1}{\sqrt{\rho_1}}\right)^4-1\right] pd^2\cdot \frac{\rho_1^2 (\rho_1+\rho_2)}{(\rho_1 + \rho_2 - 1)^3} +\OO(p^{-c}\sigma^2 )
	 \end{align}
	 for any constant $c\in (0,c_0)$.
	 \iffalse
	\begin{align*}
		& \bigabs{ L(\hat{\beta}_2^{\MTL}) - \frac{\sigma^2}{\rho_1 + \rho_2 - 1} - (1 + \frac{6}{\rho_1} + \frac{1}{\rho_1^2})pd^2\cdot \frac{\rho_1^2 (\rho_1+\rho_2)}{(\rho_1 + \rho_2 - 1)^3} } \\
		\le & 	\frac{4}{\sqrt{\rho_1}}(1 + \frac{1}{\rho_1})	pd^2\cdot \frac{\rho_1^2 (\rho_1+\rho_2)}{(\rho_1 + \rho_2 - 1)^3} + \oo(\sigma^2 + pd^2).
%	 \frac{\sigma^2}{\rho_1+\rho_2-1}+\left( 1-\frac{1}{\sqrt{\rho_1}}\right)^4  +\oo(\sigma^2 + pd^2)  \le \\
%	 \le  \frac{\sigma^2}{\rho_1+\rho_2-1}+\left( 1+\frac{1}{\sqrt{\rho_1}}\right)^4 pd^2\cdot \frac{\rho_1^2 (\rho_1+\rho_2)}{(\rho_1 + \rho_2 - 1)^3} +\oo(\sigma^2 + pd^2)
	 \end{align*}
	 \fi
\end{proposition}

@@

In classical Rademacher or VC based theory of multi-task learning, the generalization bounds are usually presented for settings where the sample sizes are equal for all tasks \cite{B00,M06,MPR16}.
%More generally, such results are still applicable when all task data are being added simultaneously.
On the other hand, uneven sample sizes between different tasks (or even dominating tasks) have been empirically observed as a cause of negative transfer \cite{YKGLHF20}.
For such settings, we have also observed that adding more labeled data from one task does not always help.
Our theory accurately predicts a curious phenomenon, where increasing the sample size of the source task results in negative transfer!
%On the other hand, we have observed that adding more labeled data does not always improve performance in multi-task learning.

@@

We apply the result to the setting of Figure \ref{fig_size} (described in Appendix \ref{app_synthetic}).
There are several unexplained observations near $y = 0$ caused by $\nu$.
The proof of Proposition \ref{prop_data_size} can be found in Appendix \ref{app_proof_32}.

\subsection{Covariate Shift} \label{sec_covshift}
So far we have considered the isotropic model where $\Sigma_1 = \Sigma_2$.
This setting is relevant for settings where different tasks share the same input features such as multi-class image classification.
In general, the covariance matrices of the two tasks may be different such as in text classification.
In this part, we consider what happens when $\Sigma_1 \neq \Sigma_2$.
We show that when $n_1 / n_2$ is large, MTL with covariate shift can be suboptimal compared to MTL without covariate shift.

%\noindent\textit{Example.}

\begin{example}
	We measure covariate shift by the transfer matrix $M = \Sigma_1^{1/2} \Sigma_2^{-1/2}$. %similar to Theorem \ref{thm_main_informal}.
	We assume that the two tasks have the same ground truth model parameters, that is, $\beta_1 = \beta_2$.
	We compare two cases: (i) when $M = \id_{p\times p}$; (ii) when $M$ has $p/2$ singular values that are equal to $\lambda$ and $p/2$ singular values that are equal to $1 / \lambda$.
	Hence, $\lambda$ measures the severity of the covariate shift.
	Figure \ref{fig_covariate} shows a simulation of this setting by varying $\lambda$.
	We observe that as source/target sample ratio increases, the performance gap between the two cases increases.
	\end{example}

%By applying Lemma \ref{lem_cov_shift_informal}, we find that when $n_1 / n_2$ is large, having no covariate shift is the optimal choice provided that the determinant of $M^{\top}M$ is bounded.
We compare different choices of $M$ that belong to the following bounded set.
Let $\lambda_i$ be the $i$-th singular value of $M$.
Let $\mu_{\min} < \mu < \mu_{\max}$ be fixed values that do not grow with $p$.
{\begin{align*}
		\cS_{\mu}\define\bigset{M \left| \prod_{i=1}^p \lambda_i \le \mu^p, \mu_{\min} \le \lambda_i\le \mu_{\max}, \text{ for all } 1\le i\le p\right.}.
\end{align*}}
%	We assume that $\beta_1$ and $\beta_2$ are generated following the isotropic model with $d = 0$.
\begin{proposition}[Covariate shift]\label{prop_covariate}
	For the setting of the example above, assume that $\beta_1 = \beta_2$, $\rho_1$ and $\rho_2$ are both greater than one.
	Let $g(M)$ denote the prediction loss of $\hat{\beta}_t^{\MTL}$ when $M = \Sigma_1^{1/2}\Sigma_2^{-1/2} \in\cS_{\mu}$.
	We have that
		\[ g(\mu\id) \le \bigbrace{1+ \bigo{{\rho_2}/{\rho_1}  }} \min_{M\in\cS_{\mu}} g(M). \]
\end{proposition}

@@

observed in Taskonomy \cite{ZSSGM18}.

>>>>

%We provide precise explanations to the phenomena of negative transfer in multi-task learning.
%We provide tight bounds on the bias and variance of the multi-task estimator for two tasks.
%The results in Section \ref{sec_general} show when multi-task learning outperforms single-task learning for general task covariance matrices and ground truth model parameters.
%In particular, the general conditions depend on specific properties of task data such as the covariate shift matrix.
%However, the conditions given by the limiting bias and variance of multi-task learning can be difficult to interpret.
%The goal of this section is to address this issue.
%We interpret the limiting bias and variance equations in an isotropic covariance setting that is described below and a simplified covariate-shifted setting.
%We show that three properties of task data affect the performance of multi-task learning as follows.
%(i) \textit{Task similarity}: We explain the phenomenon of negative transfer precisely as tasks become more different.
%(ii) \textit{Sample ratio}: We further explain a curious phenomenon where increasing the sample ratio between the source and target task helps initially, but hurts eventually.
%(iii) \textit{Covariate shift}: Finally, we show that as the sample size of the source task increases, we show that the covariate shift worsens the performance of the multi-task estimator.
%Finally, we provide a extend our results from two tasks to many tasks with the same features.
%We explain from three perspectives, including \textit{task similarity}, \textit{sample size} and \textit{covariate shift}.
%We show how negative transfer occurs by varying task similarity or sample size.
%Then we show that when source task sample size becomes large, covariate shift causes more negative effects.
We discuss several examples to illustrate our results.
Through these examples, we provide generalization bounds of hard parameter sharing estimators that depend explicitly on the sample sizes and the covariance matrices of the tasks.
We use our generalization bounds to interpret several empirical phenomena that arises from applying hard parameter sharing in practice.
For our examples, we assume that a random-effects model holds.
We assume every $\beta_i$ consists of two random components, one that is shared among all tasks and one that is task-specific.
Thus, each task contributes a certain amount to the shared component and injects a task-specific bias.
%	The labels are $Y_i = X_i\beta_i + \varepsilon_i$, where $\e_i$ consists of i.i.d. entries with mean zero and variance $\sigma^2$.
%	For our purpose, it is enough to think of the order of $d$ being $1/\sqrt{p}$ and $pd^2/\sigma^2$ being constant.
%The more precise conditions on the relations between $d^2$, $\sigma^2$ and $\kappa^2$ are given in  \eqref{choiceofpara}.
%	We assume that all the random variables have subexponential decay, while keeping in mind that our results can be applied under weaker moments assumptions as shown in Appendix \ref{sec_maintools}.

@@

We begin by considering a setting where the feature covariates are the same across all tasks but their labels are different.
In the random-effects model, we assume that every $\beta_i$ is equal to a shared component $\beta_0$ plus a task-specific component.
Suppose that the entries of $\beta_0$ are sampled i.i.d. from an isotropic Gaussian distribution.
Suppose further that $\beta_i$ is equal to $\beta_0$ plus a random Gaussian vector whose coordinates all have mean zero and variance $p^{-1} d^2 / 2$.
Thus, for any two $\beta_i$ and $\beta_j$, their distance is roughly $d^2$.
Using our results, we show that the prediction loss of the hard parameter sharing estimator is as follows.

\begin{corollary}[Derived from Theorem \ref{thm_many_tasks}]\label{prop_same_cov}
	In the setting described above, for any task $i$, the prediction loss of $\hat{\beta}_i^{\MTL}$ satisfies that
	\[ L(\hat{\beta}_i^{\MTL}). \]
\end{corollary}

@@

The above setting assumes that the feature covariates are the same across all tasks.
Hence, the sample sizes are also the same among all tasks.
Next, we assume that the feature covariates are different and consider what happens the sample sizes are different for different tasks.
We consider a two-task case where both tasks have isotropic covariance matrices.
%	Each task has sample size $n_1 = \rho_1 \cdot p$ and $n_2 = \rho_2 \cdot p$.
	%And $X_1\in\real^{n_1\times p}, X_2\in\real^{n_2\times p}$ denote the covariates of the two tasks, respectively.
%	For the ground truth linear model, we consider a random-effects model where the two tasks differ by an independent Gaussian vector.
%	Specially, for task two, we assume that its model parameter, denoted by $\beta_2$, consists of i.i.d. Gaussian entries with mean zero and variance $\kappa^2$.
%	For task one, we assume that its model parameter, denoted by $\beta_1$, is equal to $\beta_2$ plus i.i.d. Gaussian entries with mean $0$ and variance $d^2$.
%	Therefore, the two tasks have different ground truth models. In particular, the distance between $\beta_1$ and $\beta_2$ roughly scales as $d^2 \cdot p$.
%\begin{example}[Varying model distance]
%\end{example}
Using Theorem \ref{thm_main_RMT}, we derive the prediction loss of the hard parameter sharing estimator.
Without loss of generality, we describe the result for task two and remark that a similar result for task one holds.

\begin{corollary}[Derived from Theorem \ref{thm_main_RMT}]\label{cor_MTL_loss}
	In the setting described above, assume that there exists a small constant $c_0>0$ such that the signal-to-noise ratios satisfy that
	\be\label{choiceofpara0}
		p^{ c_0} \le \frac{p\kappa^2}{\sigma^2 }  \le p^{1-c_0},\quad  \frac{pd^2}{\sigma^2}=\OO(1).
	\ee
	%Let $B, A_1, A_2$ be the global minimum of equation \eqref{eq_mtl}. Then
	The prediction loss of the hard parameter sharing estimator for task two satisfies that
	\begin{align}\label{cor_MTL_error}
	%-\left[1- \left( 1-\frac{1}{\sqrt{\rho_1}}\right)^4\right] pd^2\cdot \frac{\rho_1^2 (\rho_1+\rho_2)}{(\rho_1 + \rho_2 - 1)^3} +\OO(p^{-c}\sigma^2)  \le 
	 & \left|L(\hat{\beta}_2^{\MTL}) -\frac{\sigma^2}{\rho_1+\rho_2-1} - pd^2\cdot \frac{\rho_1^2 (\rho_1+\rho_2)}{(\rho_1 + \rho_2 - 1)^3}\right|  \le  \left[\left( 1+\frac{1}{\sqrt{\rho_1}}\right)^4-1\right] pd^2\cdot \frac{\rho_1^2 (\rho_1+\rho_2)}{(\rho_1 + \rho_2 - 1)^3} +\OO(p^{-c_{\varphi}}\sigma^2 ).
	 \end{align}
\end{corollary}

@@

%In classical Rademacher or VC based theory of multi-task learning, the generalization bounds are usually presented for settings where the sample sizes are equal for all tasks \cite{B00,M06,MPR16}.
%More generally, such results are still applicable when all task data are being added simultaneously.
%On the other hand, uneven sample sizes between different tasks (or even dominating tasks) have been empirically observed as a cause of negative transfer \cite{YKGLHF20}.
%For such settings, we have also observed that adding more labeled data from one task does not always help.
%Our theory accurately predicts a curious phenomenon, where increasing the sample size of the source task results in negative transfer!
%On the other hand, we have observed that adding more labeled data does not always improve performance in multi-task learning.

@@

%We apply the result to the setting of Figure \ref{fig_size} (described in Appendix \ref{app_synthetic}).
%There are several unexplained observations near $y = 0$ caused by $\nu$.
%The proof of Proposition \ref{prop_data_size} can be found in Appendix \ref{app_proof_32}.

\subsection{Covariate Shift} \label{sec_covshift}
%So far we have considered the isotropic model where $\Sigma_1 = \Sigma_2$.
%This setting is relevant for settings where different tasks share the same input features such as multi-class image classification.
%In general, the covariance matrices of the two tasks may be different such as in text classification.

In our final example, we consider what happens when the covariance matrices are also different between different tasks, which are often called ``covariate shift''.
Our results apply to arbitrary ``covariate shift'' and we consider a simple setting of two tasks.
As we are going to show later, covariate shift is accurately captured by the spectrum of $\Sigma^{1/2}\Sigma^{-1/2}$.
To illustrate our result, we assume that the first half of the singular values of $\Sigma^{1/2}\Sigma^{-1/2}$ are all equal to a parameter $\lambda$ and the rest are all equal to one.
Our results imply that the prediction loss of the hard parameter sharing estimator is given as follows.
%We compare two cases: (i) when $M = \id_{p\times p}$; (ii) when $M$ has $p/2$ singular values that are equal to $\lambda$ and $p/2$ singular values that are equal to $1 / \lambda$.
%Hence, $\lambda$ measures the severity of the covariate shift.
%	Figure \ref{fig_covariate} shows a simulation of this setting by varying $\lambda$.
%	We observe that as source/target sample ratio increases, the performance gap between the two cases increases.
%	\end{example}

%By applying Lemma \ref{lem_cov_shift_informal}, we find that when $n_1 / n_2$ is large, having no covariate shift is the optimal choice provided that the determinant of $M^{\top}M$ is bounded.
%We compare different choices of $M$ that belong to the following bounded set.
%Let $\lambda_i$ be the $i$-th singular value of $M$.
%Let $\mu_{\min} < \mu < \mu_{\max}$ be fixed values that do not grow with $p$.
%{\begin{align*}
%		\cS_{\mu}\define\bigset{M \left| \prod_{i=1}^p \lambda_i \le \mu^p, \mu_{\min} \le \lambda_i\le \mu_{\max}, \text{ for all } 1\le i\le p\right.}.
%\end{align*}}
%	We assume that $\beta_1$ and $\beta_2$ are generated following the isotropic model with $d = 0$.
\begin{corollary}[Derived from Theorem \ref{thm_main_RMT}]\label{prop_covariate}
	In the setting described above, assume further that a random-effects model holds for both tasks.
	The prediction loss of $\hat{\beta}_2^{\MTL}$ for task two satisfies that
	\begin{align*}
		L(\hat{\beta}_2^{\MTL})
	\end{align*}
%	Let $g(M)$ denote the prediction loss of $\hat{\beta}_t^{\MTL}$ when $M = \Sigma_1^{1/2}\Sigma_2^{-1/2} \in\cS_{\mu}$.
%	We have that
%		\[ g(\mu\id) \le \bigbrace{1+ \bigo{{\rho_2}/{\rho_1}  }} \min_{M\in\cS_{\mu}} g(M). \]
\end{corollary}

@@

### 2020-09-27 arxiv/setup.tex
wip

\subsection{Notations}
We shall use $\oo(1)$ to mean a small positive quantity that converges to 0 as $p\to \infty$. 
We shall say an event $\Xi$ holds with high probability (w.h.p.) if $\P(\Xi)\ge 1- \oo(1)$, while we say $\Xi$ holds with overwhelming probability (w.o.p.) if for any large constant $D>0$, $\P(\Xi)\ge 1- p^{-D}$ for large enough $p$. For a matrix $X$, let $\lambda_{\min}(X)$ denote its smallest singular value and $\norm{X}$ denote its spectral norm.
We shall refer to random matrices of the form $\frac {X^\top X} n$ as sample covariance matrices following the standard notations in high-dimensional statistics.

@@

We introduce several standard assumptions on the random matrices we study.
\begin{assumption}\label{assume_rm}
	Let $\varphi > 4$ be a fixed value.
	Let $\tau > 1$ be a small enough constant.
	Let $X = Z \Sigma^{1/2} \in\real^{n\times p}$ be a random matrix where $Z \in \real^{n\times p}$ consists of i.i.d. entries with zero mean and unit variance and $\Sigma \in \real^{p\times p}$ is a positive semidefinite matrix.
	We assume that $\Sigma$ and $Z$ satisfy the following (standard) conditions, respectively.
	\begin{enumerate}
		\item For every entry of $Z$, we assume that its $c$-th moment exists, that is, there exist a fixed constant $C > 0$ such that
			\begin{align}\label{assmAhigh}
				\ex{\abs{Z_{i,j}}^{\varphi}} \le C, \text{ for any } 1\le i \le n \text{ and } 1\le j \le p.
			\end{align}
		\item The sample size $n$ grows to infinity proportionally with the dimension $p$, i.e. $n / p \rightarrow \rho \in (0, 1/\tau)$ as $p$ goes infinity.
	\end{enumerate}
\end{assumption}

@@

\begin{fact}\label{lem_minv}
	Let $c_{\varphi}$ be any fixed value within $(0, \frac{\varphi - 4}{2\varphi})$.
	%Let $X  \in \real^{n\times p}$ be a random matrix that satisfies Assumption \ref{assume_rm}.
	%Let $\Sigma\in\real^{p\times p}$ denote the population covariance matrix of $X$.
	With high probability over the randomness of $X$, we have that
	\begin{enumerate}
		\item The sample covariance matrix $\frac{X^{\top}X}{n}$ is full rank.
		\item $\bigtr{\Sigma (X^{\top} X)^{-1}} = \frac{p}{n - p} + \OO(p^{-c})$, cf. Theorem 2.4 in \citet{isotropic}.
		\item The singular values of $Z^{\top}Z$ are greater than $(\sqrt{n} - \sqrt{p})^2 - n \cdot p^{-c}$ and less than $(\sqrt{n} + \sqrt{p})^2 + n \cdot p^{-c}$, cf. \citet[Theorem 2.10]{isotropic} and \citet[Lemma 3.12]{DY}.
	\end{enumerate}
\end{fact}

>>>>

\subsection{Notations}
We shall use $\oo(1)$ to mean a small positive quantity that converges to 0 as $p$ goes to infinity.
We shall say an event $\Xi$ holds with high probability (w.h.p.) if $\P(\Xi)$ goes to $1$ as $p$ goes to infinity.
%, while we say $\Xi$ holds with overwhelming probability (w.o.p.) if for any large constant $D>0$, $\P(\Xi)\ge 1- p^{-D}$ for large enough $p$.
For a matrix $X$, let $\lambda_{\min}(X)$ denote its smallest singular value and $\norm{X}$ denote its spectral norm.
We shall refer to random matrices of the form $\frac {X^\top X} n$ as sample covariance matrices following the standard notations in high-dimensional statistics.

@@

We introduce several standard assumptions on the random matrices we study.
\begin{assumption}\label{assume_rm}
	Let $\varphi > 4$ be a fixed value.
	Let $\tau > 1$ be a small enough constant.
	Let $X = Z \Sigma^{1/2} \in\real^{n\times p}$ be a random matrix where $Z \in \real^{n\times p}$ consists of i.i.d. entries with zero mean and unit variance and $\Sigma \in \real^{p\times p}$ is a positive semidefinite matrix.
	We assume the following standard conditions.
	\begin{enumerate}
		\item For every entry of $Z$, we assume that its $\varphi$-th moment exists, that is, there exist a fixed constant $C > 0$ such that
			\begin{align}\label{assmAhigh}
				\ex{\abs{Z_{i,j}}^{\varphi}} \le C, \text{ for any } 1\le i \le n \text{ and } 1\le j \le p.
			\end{align}
		\item The sample size $n$ grows to infinity proportionally with the dimension $p$, i.e. $n / p \rightarrow \rho \in (0, 1/\tau)$ as $p$ goes infinity.
	\end{enumerate}
\end{assumption}

@@

\begin{fact}\label{lem_minv}
	Let $c_{\varphi}$ be any fixed value within $(0, \frac{\varphi - 4}{2\varphi})$.
	%Let $X  \in \real^{n\times p}$ be a random matrix that satisfies Assumption \ref{assume_rm}.
	%Let $\Sigma\in\real^{p\times p}$ denote the population covariance matrix of $X$.
	With high probability over the randomness of $X$, we have that
	\begin{enumerate}
		\item When $n / p$ converges to $\rho > 1$, the sample covariance matrix $\frac{X^{\top}X}{n}$ is full rank.
		\item $\bigtr{\Sigma (X^{\top} X)^{-1}} = \frac{p}{n - p} + \OO(p^{-c_{\varphi}})$, cf. Theorem 2.4 in \citet{isotropic}.
		\item The singular values of $Z^{\top}Z$ are greater than $(\sqrt{n} - \sqrt{p})^2 - n \cdot p^{-c_{\varphi}}$ and less than $(\sqrt{n} + \sqrt{p})^2 + n \cdot p^{-c_{\varphi}}$, cf. \citet[Theorem 2.10]{isotropic} and \citet[Lemma 3.12]{DY}.
	\end{enumerate}
\end{fact}
